What can GitHub and Googlecode Issues tell us about Exceptions in Android?

### 2014-08-30 android-stacks.tex
study description

  \item  It performs a large scale post morten analysis of exception stack traces of Android projects.

@@

The ramaining of this paper is organized as follows. Section 2 presents a
background on Android Plaform and Java exception handling mechanism. 
Section 3 presents the study design. Section 4 reports study findings. 
Section 5 provides further discussions and insights.
Section 6 presents the threats to validity associated to this study. Finally Section
7 describes the related work, and Section 8 presents our conclusions and
directions for future work.

@@

The Android Application Framework supports an event-driven development model. 
Among the basic elements that comprise an Android application, we can cite the
Activity. Each application usually defines one Activity per application Window.
 Each Activity defines a set of hook methods  (e.g., OnCreate(), onPause()) that are dispatched
automatically by the Android framework (by inversion of
control), in response to infrastructure and application events.

Due to the high demand for responsiveness in mobile environment (mobile apps can easily become 
unresponsivedue due to resurce limitation and excessive network access) the
Android Application Framework also provides APIs to support concurrent 
developement (e.g. AsyncTask).

@@

\subsection{Exception Types in Java} \label{sec:extypes}

@@

In Java, once an exception is thrown, the JVM looks for the nearest enclosing exception handler
(in Java it is represented by the try-catch block), and unwinds the execution stack if necessary.
 To look for the handler on the invocation stack is claimed to increase software reusability, 
since the invoker of an operation can handle it in a wider context [ref-miller].

INCLUIR REF:
Miller, R.; Tripathi, A.: Issues with Exception Handling in Object-Oriented Systems. In:
ECOOP’97. LNCS, vol. 1241, Springer (1997) 85–103.

@@

\emph{IV- All exceptions explicitly thrown by a method should be documented.}
(\cite{mandrioli1992advances,gosling2000java,wirfs2006toward,bloch2008effective}).
The exceptions thrown by a method are an important part of methods interface,
and is required to use the method properly. The checked exceptions are already
part of the  methods signature, and the method caller is aware of the checked
exceptions being thrown by it. According to~\cite{bloch2008effective}, it is
also wise to document the explicitly thrown runtime exceptions\footnote{This
excludes the implicitly signaled JVM runtime exceptions related to programming
mistakes} as carefully as checked exceptions. Doing so, the clients of a method
will be aware of all the exceptions the method can throw. If the developer fails to
do follow this practice (specially when developing library code) it will be
difficult or even impossible for the caller to make effective use of such 
method~\cite{wirfs2006toward, bloch2008effective} and protect its code
against unforeseen exceptions.

These best practices guided our exploratory study described next.
Based on information extracted from stack traces (complemented by
bytecode and source-code analysis), we investigate whether stack characteristics
can reveal whether these practices have been obeyed. 
%Next section discusses the
%study procedure and the practices' misuses that emerged from this study.

@@

This study focus on the analysis of crash information (related to uncaught exceptions) of a family of Android applications
aiming at  extracting their common characteristics and looking for similar vulnerabilities on applications and underlaying platform.

In Java when an application crashes dut to an uncaught exception ... 
Exception Stach trace is usually (procurar por artigo).

The goal of this study was to extract exception stack 
traces embedded on issues of Android projects (hosted in GitHub and GoogleCode), looking for the 
common characteristics of such stack-trace and investigating whether
 they could reveal common defficiencies on the exception handling code of both the applications and the underlaying 
 framework.

Our study focused on open-source apps, since the information needed to perform our study
cannot be retrieved from commercial apps, whose issue reports and 
source code are not publicly available. The Android open-source apps were also 
target of other research studies [fse2013][ahimed].

Falar da assumption que stacj trá relacionada a crash?
This study is built on the following assumption: the stack traces reported on issues contain 
relevant information related to application crashes. Several studies and techniques have also based on this assumption (e.g.,~\cite{sinha2009fault}~\cite{dhaliwal2011classifying}~\cite{kim2013predicting}).

\subsection{Research Questions}
In the context of our study we formulated the following reseach questions:

@@

  \item[RQ1] \noindent\emph{What are the common characteristics of exception stack traces reported on issues?} Some of the common characteristics to be investigated are the 
 main causes reported on exception stack traces, the elements responsible for signaling such exceptions.

  \item[RQ2] \noindent\emph{What patterns of exception (mis)-use related to exception types and propagation emerge from the exception stack trace analysis? Based on a set of best practices
compiled in this study, looked for characteristics on stack traces that could pinpoint to practices that were not being obeyed.}

@@

The data used in our study to answer our research questions are: the issues related to each 
Android project available in each repository; the source code and manifest file of each application;
the bytecode of all exceptions defined and reused by Android platform.

@@

As mentioned before, this study amined at mining information available on GitHub issues,
more specifically, extracting and distilling the stack traces embedded on GItHub issues. 
Issues on Github are different from issues on dedicated bug tracking tools such as 
Bugzilla and Jira. The most important difference is that there are no predefined fields
  (e.g. severity and priority). Instead, Github uses a more open ended tagging system, on which
repositories are offered a pre-defined set of labels, but repository owners can modify 
them at will. Therefore, an issue may have none ore an arbitrary set of labels depending 
on which repository it was created. Therefore, this study is based on the assumption 
if an stack trace information embedded in an issue, it contain relevant information
 concerning the exception use -  regardless the lables associated to the issue.

@@

RESUMIR:
This study used the dataset provided by the GHTorrent project~\cite{Gousi13}, an off-line mirror of the data 
offered through the Github API.  GHTorrent has been collecting data from Git since 
February 2012. Up to Dec 2013, and when we queried GHTorrent, it included 332,864 non-fork Java
repositories \footnote {A Java project is specified in GitHub with a
specific label.}  from which only 44,323 have contained at least issue report in their lifetime. From this set only 16,837 were watched by external users. To ensure that our work targets projects that are openly used by
users other than their developers, we focused our analysis on just those projects.

From this set of 16.836 Java projects which overall contained 356.057 issues, we retrieved the metadata related to each Java project (e.g., repository's names and short descriptions), and the content of every issue and issue comment and stored in a relational database. Using the heuristics presented in Section~\ref{sec:android}, we selected 482 repositories featuring Android projects on which we perform manual inspection was performed.

As mentioned before, in this study we selected a subset of Java projects to
be analyzed in more detail. The subset chosen was comprised by the Android apps
included on GitHub (until 23 Feb 2013).
To identify Android projects, we performed a case insensitive search for the
term \textsf{android} in the repository's names and short descriptions.  The
heuristic filtered 2.542 repositories, from which 589 apps had at least one
issue containing a stack trace.

Then we performed further cleanup, inspecting the site of every Android
reporting at least one stack trace, to make sure that they represented real
mobile apps. During this clean up 106 apps were removed because they were either
example projects (i.e., toy projects) or tools to support Android development
(e.g. \textsf{selendroid}, \textsf{roboeletric} --- tools to support the testing of Android apps).
The filtered set consisted of 482 apps, from which approximately 50\% are also
hosted in Google Market Place.

This set of 482 projects contained overall 32,582 issues from which 4,208 stacktraces 
were extracted. From this set of issues 44,9\% were labled as defect.

@@

Until January 2014 there were 788 android projects available in Google code repository from which 724 defined at least 1 issue and 183 projects 
had at least one issue including a stack trace. To identify the Android projects on Googlecode we used the search facilities provided looking for 
projects mentioning "Android" on the name of the project or its description. The repository page of each project was inspected, and from this set
142 corresponded to Android applications.

To mine the information available on such Google code projects we implemented a web crawler which navegated on each project and issue page
and extracted the content of each issue to store on relational database. \footnote{Google code previoulsly provided a webservice to extract information about each project, but this service was removed in what Google called "clean-up action" which happened in June 2013.}

@@

To extract exceptions from issues, we implemented ExceptionMiner, a modular
mining tool able to connect to various repositories (such as Google Code,
GHTorrent or even directly to Bugzilla), extract issues, mine stack traces from
them and classify them in predefined categories. The main components of
ExceptionMiner are as follows:

\noindent\emph{Exception StackTrace Miner} The first step in the process is
mining references to exceptions and stack traces embedded in issues, d
GitHub Commector, and GoogleCode Issue Crawler.

@@

\noindent\emph{Exception Classifier} The next step in the ExceptionMiner process
is classifying exceptions as either checked exceptions, runtime exceptions or 
errors. The exception classifier is based on bytecode analysis (based on Design
Wizard~\cite{Brunet09}) that walks up the type hierarchy of a Java exception
until it reaches a base exception type.

\noindent\emph{Exception Signaler Classifier} After its initialization it
classifies each signaler according to its origin (i.e. Android Plaform,
Aplication, Library, Libcore, or Java) based on pattern matching between the
signaler name and the packages associated to each category.

@@

\subsubsection{Data Filtering and Manual Inspections}

COMPLETE HERE.....

@@

%\begin{figure*}
%\centering
%\includegraphics[width=\hsize]{overview.pdf}
%\caption{Overview of the study based on mixed approaches.}
%\label{fig:overviewfig}
%\end{figure*}

%The GHTorrent dataset covers a broad range of
% development activities on Github, including the all the issues per project. Such data 
%is stored in unprocessed format, in a MongoDB database, while metadata is extracted 
%and stored in a relational database.

\textbf{Step1.} To obtain the Dataset to be analized, we queried GHTorrent as detailed in Section XXX. From GHTorrent we retrieved the metadata related to each Java project (e.g., 
repository's names and short descriptions), and the content of every issue and issue comment 
and stored in a relational database.

\textbf{Step 2.} Extracting Exception Information from Issues. In this step, we used the ExceptionMiner tool,
 developed in this work, to extract the exception names and exception stack traces of each
 issue and issue comment - analyzing the information of each project at a time. This initial 
set of analyzed Java projects included libraries, web-based, Android and stand-alone applications.

\textbf{Step 3.} Restricting the Dataset to Android Projects. We select a subset of all Android 
projects available on the dataset to perform a deeper analysis (which combined mined
 information extrated from issues with source-code analysis and manual inspection 
of projects resources described on next steps). This decision was driven by the
fact that most open source Anedroid applications are known for dealing
with several sources of exceptions due to multi-threaded execution and
interaction with input/output resources; moreover they share a common
well-defined platform and a limited set of libraries.
To identify Android projects, we performed a case insensitive search for 
the term \textsf{android} in the repository's names and short descriptions.  The
heuristic filtered 2.542 repositories, from which 589 apps had at least one
issue containing a stack trace. Then we performed further cleanup, inspecting the site of every Android
reporting at least one stack trace, to make sure that they represented real
mobile apps. During this clean up 106 apps were removed because they were either
example projects (i.e., toy projects) or tools to support Android development
(e.g. \textsf{selendroid}, \textsf{roboeletric} --- tools to support the testing of Android apps).
The filtered set consisted of 482 apps, from which approximately 50\% are also
hosted in Google Market Place.

To support a deeper investigation of the stack traces of Android apps, every
exception defined on a stack trace was then classified according to its type
(e.g. Error, Exception or RuntimeException) and to its signaler. To do so, the following 
steps were performed:

>>>>

  \item  It performs a large scale post morten analysis of exception stack traces reported on Android projects hosted in GitHub and Googlecode.

@@

The ramaining of this paper is organized as follows. Section 2 presents a
background on Android Plaform and Java exception model. 
Section 3 presents the study design. Section 4 reports study findings. 
Section 5 provides further discussions and insights.
Section 6 presents the threats to validity associated to this study. Finally Section
7 describes the related work, and Section 8 concludes the paper and outlines
directions for future work.

@@

The Android Application Framework supports an event-driven development model. 
Among the basic elements that comprise an Android application, we can cite the
Activity. Each application usually defines one Activity per application window.
 Each Activity defines a set of hook methods  (e.g., OnCreate(), onPause()) that are dispatched
automatically by the Android framework (by inversion of
control), in response to infrastructure and application events.

The Android Application Framework also provides several supporting APIs, such as 
the APIs to support concurrent developement (e.g. AsyncTask). 
Such APIs particularly useful due to the high demand for responsiveness 
in mobile environment (mobile apps can easily become 
unresponsivedue due to resurce limitation and excessive network access).

@@

\subsection{Exception Model in Java} \label{sec:extypes}

@@

In Java, once an exception is thrown, the JVM looks for the nearest enclosing exception handler
(in Java it is represented by the try-catch block), and unwinds the execution stack if necessary.
 To look for the handler on the invocation stack is claimed to increase software reusability, 
since the invoker of an operation can handle it in a wider context ~/cite{miller1997issues}.

@@

\emph{IV- All exceptions explicitly thrown by a method should be documented.}
(\cite{mandrioli1992advances,gosling2000java,wirfs2006toward,bloch2008effective}).
The exceptions thrown by a method are an important part of methods interface,
and is required to use the method properly. The checked exceptions are already
part of the  methods signature, and the method caller is aware of the checked
exceptions being thrown by it. According to~\cite{bloch2008effective}, it is
also wise to document the explicitly thrown runtime exceptions\footnote{This
excludes the implicitly signaled JVM runtime exceptions related to programming
mistakes} as carefully as checked exceptions. Doing so, the clients of a method
will be aware of all the exceptions the method can throw and can protect its code
against unforeseen exceptions.

If the developer fails to do follow such practices (specially when developing 
library/reusable code) it will be difficult or even impossible for the caller to 
make effective use of such method~\cite{wirfs2006toward, bloch2008effective},
and protect the system against the uncaught exceptions - one of the main causes of 
systems crashes.

These best practices guided our exploratory study described next.
Based on information extracted from stack traces (complemented by
bytecode and source-code analysis), we investigate whether stack characteristics
can reveal whether these practices have been obeyed.

%Next section discusses the
%study procedure and the practices' misuses that emerged from this study.

@@

%This study focus on the analysis of crash information (related to uncaught exceptions) of a family of Android applications
%aiming at  extracting their common characteristics and looking for similar vulnerabilities on applications and underlaying platform.
%In Java when an application crashes dut to an uncaught exception ... 
%Exception Stach trace is usually (procurar por artigo).

The goal of this study was to extract exception stack traces embedded on issues of Android projects 
(hosted in GitHub and GoogleCode), looking for its 
common characteristics and investigating whether
 they could reveal common fragilities on the exception code of both the 
applications and the underlaying  framework.

Our study focused on open-source apps, since the information needed to perform our study
cannot be retrieved from commercial apps, whose issue reports and 
source code are not publicly available. The Android open-source apps were also 
target of other research studies ~\cite{Linar13,ahimed}.

% Falar da assumption que stacj trá relacionada a crash?
%This study is built on the following assumption: the stack traces reported on issues contain 
%relevant information related to application crashes. Several studies and techniques have also based on this assumption %(e.g.,~\cite{sinha2009fault}~\cite{dhaliwal2011classifying}~\cite{kim2013predicting}).

%\subsection{Research Questions}
In the context of our study we formulated the following reseach questions:

@@

  \item[RQ1] \noindent\emph{What are the common characteristics of exception stack traces reported on issues?} 
Some of the common characteristics to be investigated are the exceptions reported as the root cause on
the exception stack trace, the origin of such exceptions (e.g., the application, a library, the Android platform).

%Which fragilities on the exception-related code emerge from the exception stack trace analysis? 
%Can such characteristics reveal common fragilities on the exception code of applications or the underlying framework

  \item[RQ2] \noindent\emph{Which patterns of exception (mis)-use emerge?}
Based on a set of best practices compiled in this study, the study investigates whether characteristics on exception stack traces  
can pinpoint best practices violations. Chraracteristics of exception stack traces and detected violations can reveal fragilities
 on the exception handling code, which threats the application dependability. }

@@

To answer the research questions the following information is needed: (i) the issues of reported on every Android project hosted on 
GitHub and Googlecode; (ii) the stack traces embedded on those issues; (iii) the type of the exceptions
 reported on the stack traces (e.g., error, runtime, checked); (iv) the origin of such exceptions 
(e.g., the application, a library, the Android platform). Next sections describes how such information 
was retrieved from GitHub and Googlecode, how the exceptions stack traces were extracted and distilled,
 and how the type associated to each exception and its origens was identified.

%mentioned 
%The data used in our study to answer our research questions are: the issues related to each 
%Android project available in each repository; the source code and manifest file of each application;
%the bytecode of all exceptions defined and reused by Android platform.

@@

%As mentioned before, this study amined at mining information available on GitHub issues,
%more specifically, extracting and distilling the stack traces embedded on GItHub issues.

This study used the dataset provided by the GHTorrent project~\cite{Gousi13}, 
an off-line mirror of the data  offered through the Github API.  
To identify Android projects, we performed a case insensitive search for the
term \textsf{android} in the repository's names and short descriptions.  
Up to 23 Feb 2014,  when we queried GHTorrent, such heuristic could
 find 2.542 repositories,  from which 589 apps had at least one
issue containing a stack trace.

Then we performed further cleanup, inspecting the site of every Android
reporting at least one stack trace, to make sure that they represented real
mobile apps. During this clean up 106 apps were removed because they were either
example projects (i.e., toy projects) or tools to support Android development
(e.g. \textsf{selendroid}, \textsf{roboeletric} --- tools to support the testing of Android apps).
The filtered set consisted of 482 apps.

This set of 482 projects contained overall 32,582 issues from which 4,208 stacktraces 
were extracted. Issues on Github are different from issues on dedicated bug tracking tools such as 
Bugzilla and Jira. The most important difference is that there are no predefined fields
  (e.g. severity and priority). Instead, Github uses a more open ended tagging system, on which
repositories are offered a pre-defined set of labels, but repository owners can modify 
them at will. Therefore, an issue may have none ore an arbitrary set of labels depending 
on which repository it was created. Therefore, this study is based on the assumption 
if an stack trace information embedded in an issue, it contain relevant information
that can reveal violation to best practices - regardless the lables associated to the issue.

@@

@@

Different from GitHub, Googlecode does not provide an API to access the information related
 to hosted projects ~\footnote{Googlecode previoulsly provided a webservice but was deacrivated in June 2013.}.
However many well used open source Android application such as K9mail, is hosted in Googlecode which stores
the issue history of this and other apps are store there ~\footnote{K9mail moved to Github but advice their users to create issues on Googlecode as a way of not loosing the issue history.}.

To overcome this limitation and mine the information available on such Googlecode 
projects we implemented a Web crawler which navegated (incorporated in Exception Miner tool described next) 
 on each project and issue page, extracting the content of each issue to store on relational database. \footnote{Google code previoulsly provided a webservice to extract information about each project, but this service was removed in what Google called "clean-up action" which happened in June 2013.}

To overcome this limitation we needed to implement a Web Crawler (incorporated in Exception Miner tool described next) that navegated 
through the web interface of Googlecode projects extracting all issues and issue comments and storing in a relational database for later analysis.

To identify Android projects in Googlecode, we performed a similar heuristic, we performed a case insensitive search 
(on Googlecode search interface) for the term \textsf{android}.  On January 2014, when we queried Googlecode, such heuristic could
 find 788  projects, from which 724 defined at least 1 issue and, 183 projects 
had at least one issue including a stack trace. Then we performed further cleanup, inspecting the site of every Android
reporting at least one stack trace, and as a result we could identify 158 
Android projects.

@@

To support this study we developed ExceptionMiner, a modular mining tool able 
to connect to various repositories (such as Google Code,
GHTorrent or even directly to Bugzilla), extract issues, mine stack traces from
them and distill exteption stack trace information. The main components of
ExceptionMiner are as follows:

\noindent\emph{Repository Connectors} The first step in the process is
mining references to exceptions and stack traces embedded in issues, d
GitHub Commector, and GoogleCode Issue Crawler.

@@

\noindent\emph{Exception Classifier} To support a deeper investigation of the stack traces 
every exception defined on a stack trace need to be classified according to its type
(e.g. Error, Exception or RuntimeException). The exception classifier is based on 
bytecode analysis (based on Design Wizard~\cite{Brunet09}) that walks up the 
type hierarchy of a Java exception until it reaches a base exception type.

\noindent\emph{Exception Signaler Classifier} 
This module is responsible for classifying each signaler according 
to its origin (i.e. Android Plaform, Aplication, Library, Libcore, or Java) 
based on pattern matching between the signaler name and the packages associated to each category.
TELL THAT ALL PACKAGES COMPOSISNG APPS .... WERE RECOVED....

@@

\subsubsection{The Minig Process}

{MOSTRAR UMA FIGURA COM TODOS OS PASSOS E RESSALTAR OS PASSOS MANUAIS}
COMPLETE HERE.....

@@
